---
title: "Lab 6"
author: "Ana Menzies"
date: "08/04/2022"
output: pdf_document
---

#Briefly, write up your strategy in designing your experiment. What were you wanting to ascertain? What sources of variation did you have to consider? How did you control for these sources of variation (e.g., via randomisation, replication, or blocking)? What given the study set-up, could you not control for?

I completed the same experiment 4 times over different booders - brooders seperated (blocking?)

So that I knew results were not from different brooders

Randomised wheet and masie and copper so I know that tier is not making a difference 

I didnt ranomise diet as only 8  - so often largely skewed to eg maisze

Then did 4x treatments. Reperated across brooders but each time move treatments so begun at a different tier

control (0) 100, 200, 300

150


Produce a plot of your collected data, which you think best illustrates your question and the data you collected (along with any trends you think may exist). Your plot should be adequately labelled and include an informative caption. Your plot and caption should be stand-alone (i.e., the reader
should be able to understand what you are showing from simply the plot and caption.) 

```{r, fig.cap= "Figure 1: "}
# Call packages needed for analysis
library(readr)
library(ggplot2)
library(tidyverse)

# Read in data 
readr:read_csv()
data <- read.csv("chick_data-2022-04-07.csv")

ggplot(data_1, aes(x=copper, y=growth, colour=diet)) +
  geom_point() +
  geom_smooth(method = 'lm', se=FALSE)


#data_m <- data %>% 
#  filter(diet == "Maize") %>%
 # group_by(copper)
  
data_m <- data %>% 
  filter(diet == "Maize") %>%
  dplyr::select(c("copper", "growth"))%>%
  group_by(copper)
  

data_w <- data %>% 
  filter(diet == "Wheat") %>%
  dplyr::select(c("copper", "growth"))%>%
  group_by(copper)

ggplot(data_m,aes(x = copper, y = growth, group=copper)) + ## data & aesthetics
    geom_boxplot()  + ## geom
  ggtitle("Maize") +
  ylab("Growth") +
  xlab("Copper") 
 ## theme

ggplot(data_w,aes(x = copper, y = growth, group=copper)) + ## data & aesthetics
    geom_boxplot()  + ## geom
    scale_fill_viridis(discrete = TRUE, alpha=0.6) +
    geom_jitter(color="black", size=0.4, alpha=0.9) +
  ggtitle("Wheat") +
  ylab("Growth") +
  xlab("Copper") 
 ## theme

ggplot(data,aes(x = copper, y = growth, group=copper, colour = diet)) + ## data & aesthetics
    geom_boxplot()  + ## geom
  ggtitle("Maize") +
  ylab("Growth") +
  xlab("Copper") 
 ## theme

########
#Good one?

data_1 <- data %>% 
  dplyr::select(c("copper", "growth", "diet"))%>%
  group_by(diet)

str(data_1)

data_1$copper <- as.factor(data_1$copper)

data_1$diet <- as.factor(data_1$diet)


ggplot(data_1, aes(x=copper, y=growth, fill=diet)) +
  geom_boxplot() +
  ylab("Growth rate") + 
xlab("Copper units added") +
  theme_minimal()


#http://www.sthda.com/english/wiki/ggplot2-box-plot-quick-start-guide-r-software-and-data-visualization


```

Analyse your data, using any of the techniques we've covered in the course so far (or anyway else you deem appropriate). Draw inference. Did you deem it necessary to include random effects in your model? What did you find? Were there any limitations in your design? What might you be able to conclude? You should report all relevant details from your analysis. In addition, in a short and concise Appendix, you should include all relevant code (printed in code chunks) so that your analysis can be reproduced (assuming a data structure as downloaded from the application).



```{r}
lm_1 <- lm(growth ~ copper*diet, data = data)

summary(lm_1)$coef

anova(lm_1)

gglm::gglm(lm_1)

##multiple comparisons?

library(predictmeans)

data_1_pm <- data_1 %>%
  mutate(copper = as.factor(copper))

lm_2 <- lm(growth ~ copper*diet, data = data_1_pm)

pm <- predictmeans::predictmeans(lm_2 , modelterm = "copper", 
                                     pairwise = TRUE, plot = FALSE)
str(pm)
pm$mean_table
print(pm$`Pairwise p-value`)
```

as it is a balanced design, ordering doesnt matter

Knit your report to PDF and upload. Your write-up should be no more than a page of A4 in total (including the plot and code chunks).


What I will be looking for?
Legitimate experimental strategy: has an insightful and justified design been implemented and outlined?
Publication worth plot: does the plot adequately illustrate the data collected using publication-worthy aesthetics?
Warranted analysis and accurate inference: were appropriate statistical techniques used to analyse the data and were correct inferences made?
Structured and reproducible code appendix: would someone with no prior knowledge of this lab be able to reproduce your analysis?


## Appendix of code

